---
title: "SWP1 - Group 6"
author: "Thomas Siskos, Julian Cantor, Oliver Brose, Konstantin Haesler"
date: "May 3, 2018"
output:
  pdf_document: default
  html_document:
    df_print: paged
geometry: margin=2.54cm
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.align = "center", fig.height = 3, fig.width = 10)
```

```{r warning=FALSE}
library(car)
library(dplyr)
library(ggplot2)
library(GGally)
library(stargazer)
data("Salaries")
```


### Questions 1, 2, 3 & 4
There are 397 observation and 6 variables in the dataset, 21 professors with more than 40 yrs of experience and 54 professors with a salary higher than 150,000. On the other hand, the mean salary of the professors with more than 20 yrs of experience is $122,104.

```{r, eval = FALSE, echo = FALSE}
# Q1
dim(Salaries)
# Q2
old = Salaries[, "yrs.service"] > 40
sum(old)
# Q3
rich = Salaries[, "salary"] > 150000
sum(rich)
# Q4
mean(Salaries[Salaries$yrs.service > 20, "salary"])
```

### Question 5: 
To find out more of the dataset we computed some descriptive statistics.

```{r, echo = TRUE}
summary(Salaries)
```

### Question 6: 
The following chart shows the counts and proportions by rank and sex.  
```{r, echo = T}
counts = Salaries %>%
  count(sex, rank) %>%
  mutate(prop = n / sum(n))
print(counts)
```

### Question 7: 
The histogram below shows the desired output. 
```{r}
ggplot(Salaries, aes(x = yrs.service)) +
  geom_histogram(aes(y = ..density..), binwidth = 5) +
  geom_density(color="red") +
  theme_bw() +
  ggtitle("Distribution of years of service")
```

### Question 8: 
The bloxplot below shows the desired output. 

```{r fig.width=2, fig.align="right"}
ggplot(Salaries, aes(x = 1, y = salary)) +
  geom_boxplot() +
  theme_bw() +
  ggtitle("Boxplot salary") +
  xlab("")
```

### Question 9: 
The bloxplot below shows the desired output.
```{r}
ggplot(Salaries, aes(x = rank, y = salary, group = rank)) +
  geom_boxplot() +
  coord_flip() +
  theme_bw() +
  ggtitle("Boxplot salaries by rank")
```

### Question 10: 
The scatterplot shows the desired result. 
```{r}
ggplot(Salaries, aes(x = yrs.since.phd, y = salary)) +
  geom_point() +
  theme_bw() +
  ggtitle("Salary vs. years since Phd")
```

### Question 11: 
The correlation between salary and years since phd is 0.41: With the model output shown below one can see that the p-value is less than 0.05 and thus the correlation has statistical significance.
```{r, echo = T, results="asis"}
cor(Salaries$salary, Salaries$yrs.since.phd)
cor(Salaries$salary, Salaries$yrs.service)
model = lm("salary ~ yrs.since.phd", data = Salaries)
print(summary(model))
```

### Question 12: 
The following diagram shows all bivariate relationships.
```{r fig.height=7}
ggpairs(Salaries) +
  theme_bw()
```

### Question 13: 
The next table shows mean salaries by rank and sex. 
```{r}
s <- Salaries %>%
  group_by(rank, sex) %>%
  summarise(mean.salary = mean(salary)) 

print(xtabs(mean.salary~., data=s))
```

### Question 14: 
```{r}
ggplot(Salaries, aes(x = sex, y = salary, group = sex)) +
  geom_boxplot() +
  theme_bw() +
  coord_flip() +
  ggtitle("Distribution of salaries conditioned on sex")
```

### Question 15: 
We used a Chi-square test because it is appropiate for testing differences in proportions. We can not reject the null hipotesis of the Chi-square test since the p-value of the test is larger than 0.05. This means than we can not conclude that there is a difference between the proportions of women in each discipline. 
```{r}
cont = table(Salaries$sex, Salaries$discipline)
print(cont)
chisq.test(cont)
```
